---
title: "STA 141A: Assignment 6"
author: "(Type your name here)"
output: html_document
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Setup ---------------------------------------------------------------

set.seed(141)

library(tidyverse)
library(ISLR)         # for Default dataset
library(rpart)        # decision trees
library(rpart.plot)   # tree plotting
library(randomForest) # random forest

```

---
## Question 1

We will use the `mtcars` dataset to build a prediction model for fuel economy.

1. Treat `mpg` (miles per gallon) as the response and consider `wt` (weight) and `hp` (horsepower) as predictors. Split the data into a training set ($80\%$) and a test set ($20\%$) using a random split.
2. Fit the linear regression model
   $$ \text{mpg} = \beta_0 + \beta_1 \, \text{wt} + \beta_2 \, \text{hp} + \varepsilon $$
   using the training set.
3. Compute the training MSE and test MSE.
4. Briefly comment on the difference between training and test MSE.

## Question 2

We use the `Default` dataset from the `ISLR` package. The response `default` indicates whether a customer defaults on their credit card balance (`Yes` or `No`).

1. Convert `Default` to a tibble and create a numeric binary variable

   $$
   \texttt{default01} =
   \begin{cases}
     1 & \text{if default = "Yes"} \\
     0 & \text{if default = "No"}
   \end{cases}
   $$

2. Fit a logistic regression model with `default01` as the response and predictors `balance` and `income` using the full dataset.
3. For each observation, compute the fitted probability of default.
4. Using a classification threshold of 0.5, assign each observation to “default” (1) or “no default” (0). Compute the overall classification accuracy.
5. Now change the threshold to 0.2. Recompute the classification accuracy. Compare the accuracy under thresholds 0.5 and 0.2

## Question 3

We consider the `Default` dataset again.

1. Using the same dataset created in Problem 2, fit a classification tree to predict `default01` using`balance`, `income`, and `student`. You may use `rpart()` with `method = "class"`.
2. Plot the fitted tree with `rpart.plot()`.
3. Compute the overall classification accuracy of the tree on the full dataset.
4. Compare the tree’s accuracy to the logistic regression model from Problem 2 (using threshold 0.5). Which model classifies better in this dataset?


## Question 4

We now fit a random forest model for the same default prediction task.

1. Using the same dataset in the previous questions, fit a random forest to predict `default01` using `balance`, `income`, and `student`. Make sure the response is treated as a factor in the model.
2. Print the random forest object and note the out-of-bag (OOB) error rate.
3. Create a variable importance plot.
4. Compute the classification accuracy of the random forest on the full dataset.
5. Compare the random forest accuracy with  the logistic regression (Question 2, threshold 0.5) and  the classification tree (Question 3). 


## Question 5

We now compare three models for predicting default:

- Logistic regression (Problem 2),
- Classification tree (Problem 3),
- Random forest (Problem 4).

Use 5-fold cross-validation on the `Default` data.

1. Write a function `cv_error(model_type, k = 5, data = df_def)` that:
   - randomly splits the data into `k` folds,
   - for each fold, fits the specified model on the other `k-1` folds,
   - computes the misclassification rate on the held-out fold,
   - returns the average misclassification rate across folds.

   Here `model_type` is one of `"logit"`, `"tree"`, `"rf"`.

2. Use your function to compute 5-fold CV errors for all three models.
