---
title: "STA 141A: Assignment 3"
author: "(Type your name here)"
output: html_document
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Random seed 

```{r}
set.seed(1)
```

---

## Question 1 


Consider an exponential distribution $X \sim \text{Exponential}(\text{rate} = 1).$


For sample sizes $n = 20, 50, 200$, conduct simulation studies with 5,000 replicates as follows. 

1. Simulate the sampling distribution of the $10\%$ trimmed mean (`mean( , trim=0.1)`).
2. Plot the sampling distributions of the trimmed means for each $n$. 
3. Compute the empirical standard error of the trimmed means for each $n$.
4. Compare the standard error of the trimmed mean to that of the ordinary mean and comment on your findings, if any.


## Question 2


Consider a standard normal distribution $X \sim N(0, 1)$.

For sample sizes $n = 10, 50, 200$, conduct the simulation studies with 5000 replicates as follows. 

1. Simulate the sampling distribution of the sample variance $S^2$.
2. Plot the sampling distribution of the sample variance for each $n$. 
3. Compute the empirical means of $S^2$ and compare them to the population variance.
4. Comment on any apparent bias and on the shape of the sampling distribution.


## Question 3

Consider an exponential distribution $X \sim \text{Exponential}(1)$. Let $q$ be the 75th percentile of the population.  For $n = 10, 50, 200$, estimate $P(\bar{X} > q)$ using Monte Carlo simulation. 

## Question 4

Consider the Student $t$-distribution with 3 degrees of freedom $X \sim t(3)$. 

1. Draw a single sample of size $n = 100$.
2. Construct a *bootstrap distribution* of the sample mean using $B = 4000$ resamples.
3. Simulate the sampling distribution of the mean using the Monte Carlo method. 
4. Compare the standard error of the mean from bootstrap and Monte Carlo method.

